\section{Axiom Verification}
\label{app:axioms}

This appendix documents the axioms used by the main theorems, verifying that no custom axioms are introduced.

\subsection{Standard mathlib Axioms}

The Lean~4 mathematical library (mathlib) uses three standard axioms:

\begin{description}
\item[\texttt{propext}] \textbf{Propositional extensionality}: Two propositions that are logically equivalent are equal.
\[
(P \leftrightarrow Q) \to (P = Q)
\]
This is standard in classical mathematics and required for working with propositions as types.

\item[\texttt{Classical.choice}] \textbf{Axiom of choice}: For any nonempty type, there exists an element.
\[
\forall \alpha.\, \text{Nonempty}(\alpha) \to \alpha
\]
This is essential for measure theory, which relies heavily on the axiom of choice.

\item[\texttt{Quot.sound}] \textbf{Quotient soundness}: If two elements are related by an equivalence relation, their quotient images are equal.
\[
r\, a\, b \to \mathsf{Quot.mk}\, r\, a = \mathsf{Quot.mk}\, r\, b
\]
This is standard for quotient type constructions.
\end{description}

\subsection{Axiom Checks}

The following Lean commands verify the axioms used by each theorem:

\paragraph{Main theorem (martingale proof).}
\begin{lstlisting}
#print axioms Exchangeability.DeFinetti.deFinetti
-- 'Exchangeability.DeFinetti.deFinetti' depends on axioms:
--   [propext, Classical.choice, Quot.sound]
\end{lstlisting}

\paragraph{Full equivalence.}
\begin{lstlisting}
#print axioms Exchangeability.DeFinetti.deFinetti_RyllNardzewski_equivalence
-- 'Exchangeability.DeFinetti.deFinetti_RyllNardzewski_equivalence' depends on axioms:
--   [propext, Classical.choice, Quot.sound]
\end{lstlisting}

\paragraph{$L^2$ proof.}
\begin{lstlisting}
#print axioms Exchangeability.DeFinetti.conditionallyIID_of_contractable_viaL2
-- 'Exchangeability.DeFinetti.conditionallyIID_of_contractable_viaL2' depends on axioms:
--   [propext, Classical.choice, Quot.sound]
\end{lstlisting}

\paragraph{Koopman proof.}
\begin{lstlisting}
#print axioms Exchangeability.DeFinetti.conditionallyIID_of_contractable_viaKoopman
-- 'Exchangeability.DeFinetti.conditionallyIID_of_contractable_viaKoopman' depends on axioms:
--   [propext, Classical.choice, Quot.sound]
\end{lstlisting}

\subsection{Verification}

All theorems depend on exactly the same three standard axioms. Notably:
\begin{itemize}
\item No custom \texttt{axiom} declarations are used.
\item No \texttt{sorry} placeholders appear in any theorem.
\item No non-standard axioms (such as \texttt{Axiom.choice} variants beyond the standard one) are introduced.
\end{itemize}

\subsection{Running the Verification}

To verify axiom hygiene yourself:
\begin{verbatim}
cd exchangeability
lake env lean paper_materials/PaperIntrospection.lean
\end{verbatim}

This file contains the \texttt{\#print axioms} commands and will display the axiom dependencies when run.
