\section{Reproducibility}
\label{app:reproduce}

This appendix provides instructions for building and verifying the formalization.

\subsection{Prerequisites}

\begin{itemize}
\item \textbf{Lean 4}: Version 4.27.0-rc1 or compatible
\item \textbf{Lake}: Version 5.0.0 or compatible (included with Lean)
\item \textbf{Git}: For cloning the repository
\end{itemize}

\paragraph{Installing Lean.}
Follow the instructions at \url{https://lean-lang.org/lean4/doc/setup.html} or use elan:
\begin{verbatim}
curl https://raw.githubusercontent.com/leanprover/elan/master/elan-init.sh -sSf | sh
\end{verbatim}

\subsection{Building the Formalization}

\paragraph{Clone the repository.}
\begin{verbatim}
git clone https://github.com/cameronfreer/exchangeability
cd exchangeability
\end{verbatim}

\paragraph{Build the project.}
\begin{verbatim}
lake build
\end{verbatim}

Expected output:
\begin{verbatim}
Build completed successfully (3296 jobs).
\end{verbatim}

Build time is approximately 5 minutes on a modern machine.

\paragraph{Build a specific module.}
\begin{verbatim}
lake build Exchangeability.DeFinetti.Theorem
\end{verbatim}

\subsection{Verifying the Main Theorems}

\paragraph{Check theorem signatures.}
Create a file \texttt{check.lean}:
\begin{lstlisting}
import Exchangeability.DeFinetti.Theorem

#check @Exchangeability.DeFinetti.deFinetti
#check @Exchangeability.DeFinetti.deFinetti_equivalence
#check @Exchangeability.DeFinetti.deFinetti_RyllNardzewski_equivalence
\end{lstlisting}

Run:
\begin{verbatim}
lake env lean check.lean
\end{verbatim}

\paragraph{Verify axioms.}
\begin{verbatim}
lake env lean paper_materials/PaperIntrospection.lean
\end{verbatim}

This displays the axiom dependencies for all main theorems.

\subsection{Repository Structure}

\begin{verbatim}
exchangeability/
  lakefile.lean         -- Build configuration
  lake-manifest.json    -- Dependency versions
  Exchangeability/      -- Main source code
    DeFinetti/          -- de Finetti theorem
    Ergodic/            -- Ergodic theory
    Probability/        -- Probability infrastructure
    Tail/               -- Tail sigma-algebra
    ...
  paper_materials/      -- Paper preparation materials
  paper_draft/          -- This paper (LaTeX source)
  README.md             -- Overview documentation
  CLAUDE.md             -- Development notes
  STATUS.md             -- Completion status
\end{verbatim}

\subsection{Version Information}

The formalization was developed and tested with:
\begin{itemize}
\item \textbf{Lean}: v4.27.0-rc1
\item \textbf{Lake}: v5.0.0-src+2fcce72
\item \textbf{mathlib4}: See \texttt{lake-manifest.json} for exact commit
\end{itemize}

\subsection{Troubleshooting}

\paragraph{Cache download.}
To speed up builds, download precompiled cache:
\begin{verbatim}
lake exe cache get
\end{verbatim}

\paragraph{Clean rebuild.}
If builds fail, try a clean rebuild:
\begin{verbatim}
lake clean
lake build
\end{verbatim}

\paragraph{mathlib version issues.}
If there are version conflicts, update mathlib:
\begin{verbatim}
lake update
lake exe cache get
lake build
\end{verbatim}

\subsection{Continuous Integration}

The repository includes CI configuration that:
\begin{itemize}
\item Builds the full project on each push
\item Verifies that all theorems compile without \texttt{sorry}
\item Checks for linting warnings
\end{itemize}

Build status is displayed in the repository README.

\subsection{Contact}

For questions or issues:
\begin{itemize}
\item Open an issue at \url{https://github.com/cameronfreer/exchangeability/issues}
\item Email: [author email]
\end{itemize}
